\textit{Supporting Information for}: Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-02-17 paper/supporting_information.tex
supporting info edits

\section*{Additional details about topic modeling methods and results}
\subsection*{Optimizing topic model parameters}
In order to create accurate video and recall models, we used an optimization method that was driven by our ability to explain hand-annotated memory performance metrics collected by \cite{ChenEtal17}.  Specifically, we used a grid search to compute the $\omega$ (video sliding window duration, in scenes), $\rho$ (recall sliding window duration, in sentences), and $K$ (number of topics) that satisfied
\[
\argmax_{\omega, \rho, K} \left[\mathrm{corr}\left(\mathrm{corr}\left(\mu\left(\omega, \rho, K\right), \nu\left(\omega, \rho, K\right)\right), \theta\right)\right],
\]
where $\mathrm{corr}(\mu, \nu)$ is the per-participant correlation between the temporal correlation matrices of the video ($\mu$) and recall ($\nu$) trajectory, and $\theta$ is the per-participant hand-annotated memory performance.  We searched over a grid of pre-specified values for each of these parameters; the resulting correlations are displayed in Figure~\ref{fig:paramsearch}.  The optimal parameters were $\omega = 50$, $\rho = 10$, and $K = 100$.  In subsequent refinements of our topic modeling approach, we made various improvements to how we parse the video and recall text and construct the trajectories' timeseries, but maintained these optimized parameters throughout.

@@

\begin{figure}[]
\centering
\includegraphics[width=0.65\textwidth]{figs/topic_words}
\caption{\small \textbf{Topics discovered in \textit{Sherlock}.} We applied a topic model to hand-annotated information about 1000 scenes spanning the 45 minute episode.  We identified 27 unique topics with non-zero weights (we used $K=100$ topics to fit the model).  Each topic comprises a distribution of weights over all words in the vocabulary.  For each topic, we show the words with the 10 largest weights, along with a suggested description of the topic.}
\label{fig:topics}
\end{figure}
%\FloatBarrier

@@

\subsection*{Creating a low-dimensional embedding space}
Figures~5 and 6C in the main text display two-dimensional projections of the 100-dimensional topic trajectories for the video (5A, 6C), average recall (5B), and each individual's recall (5C, 6C).  We created these embeddings using the Uniform Manifold Approximation and Projection algorithm \citep[UMAP;][]{McInEtal18} called from our high-dimensional visualization and text analysis software, \texttt{HyperTools}~\citep{HeusEtal18a}.  An advantage of the UMAP algorithm over comparable manifold learning techniques (such as $t$-SNE) is that UMAP explicitly attempts to preserve the global structure of the data \citep{McInEtal18,BechEtal19} by constructing a space where distance on the manifold is standard Euclidean distance, with respect to the global coordinate system.  This was important in our use case, as we wanted to visualize both the evolving structure of the video and the spatial relationships between presented and recalled content.

@@

We performed a grid search over pre-specified values of these hyper-parameters, and optimized the manifold space for visualization based on two criteria: First, that the 2D embedding of the video trajectory should reflect its original 100-dimensional structure as faithfully as possible. Second, that the path traversed by the embedded video trajectory should intersect itself a minimal number of times.  The first criteria helps bolster the validity of visual intuitions about relationships between sections of video content, based on their locations in the embedding space.  The second criteria was motivated by the observed low off-diagonal values in the video trajectory's temporal correlation matrix (suggesting that the same topic-space coordinates should not be revisited; see Figure~2A in the main text) and the inability to accurately describe the consistency of participants' transitions through topic space at locations where those transitions frequently occurred in multiple, orthogonal directions (see \textit{Methods} for analysis details).  We formalized this optimization as the combination of hyper-parameter values that satisfied
\[
\argmax_{\left\{K, \gamma, \delta, \tau, \varphi~\in~E~\mathrm{|}~\Phi\left(K, \gamma, \delta, \tau, \varphi \right) \right\}} \left[\mathrm{corr}\left(\Upsilon\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right),~\Upsilon\left(\chi\right)\right)\right],~\mathrm{where}
\]
\[
\Phi\left(K, \gamma, \delta, \tau, \varphi \right) = \argmin_{K, \gamma, \delta, \tau, \varphi} \left[\Gamma\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right)\right],
\]
$\xi$ is the video's trajectory through the manifold space; $\chi$ is the original, 100-dimensional video trajectory; $\Upsilon$ is a condensed matrix of pairwise distances between event vectors (computed using correlation distance in the original topic space and Euclidean distance in the manifold space); $\mathrm{corr}\left(\Upsilon\left(\xi\right), \Upsilon\left(\chi\right)\right)$ is the correlation between the sets of pairwise distances, and $\Gamma$ is the number of instances in which lines drawn between consecutive video event embeddings intersected each other.   The sets of hyper-parameter values we searched over comprised: $K \in \{10x \in \mathbb{Z}~|~10 < x <  22\} \cup \{161\}$ (a range roughly centered on half the total number of events, 161, in order to balance representations of local and global structure); $\gamma \in \{1, 3, 5, 7, 9\}$; $\delta \in \{0.1, 0.3, 0.5, 0.7, 0.9\}$; $\tau \in \{x \in \mathbb{Z}~|~0 < x < 101\}$; and $\varphi \in {S\choose3},~\mathrm{where}~S=\{\mathrm{video~events,~average~recall~events, individual~recall~events}\}$.  The optimal parameters (that yielded $\Phi=0$) were $K=170$, $\gamma=7$, $\delta=0.7$, $\tau=41$, with the order of sequence $\varphi$ as the average recall events, video events, and individuals' recall events, vertically concatenated, in order.

>>>>

\section*{Additional details about topic modeling methods and results}
\subsection*{Optimizing topic model parameters}
In order to create accurate video and recall models, we used an optimization method that was driven by our ability to explain hand-annotated memory performance metrics collected by \cite{ChenEtal17}.  Specifically, we used a grid search to compute the $\omega$ (video sliding window duration, in scenes), $\rho$ (recall sliding window duration, in sentences), and $K$ (number of topics) that satisfied
\[
\argmax_{\omega, \rho, K} \left[\mathrm{corr}\left(\mathrm{corr}\left(\mu\left(\omega, \rho, K\right), \nu\left(\omega, \rho, K\right)\right), \theta\right)\right],
\]
where $\mathrm{corr}(\mu, \nu)$ is the per-participant correlation between the temporal correlation matrices of the video ($\mu$) and recall ($\nu$) trajectory, and $\theta$ is the per-participant hand-annotated memory performance.  We searched over a grid of pre-specified values for each of these parameters; the resulting correlations are displayed in Figure~\ref{fig:paramsearch}.  The optimal parameters were $\omega = 50$, $\rho = 10$, and $K = 100$.

@@

\begin{figure}[]
\centering
\includegraphics[width=0.65\textwidth]{figs/topic_words}
\caption{\small \textbf{Topics discovered in \textit{Sherlock}.} We applied a topic model to hand-annotated information about 1000 scenes spanning the 45 minute episode.  We identified 32 unique topics with non-zero weights (we used $K=100$ topics to fit the model).  Each topic comprises a distribution of weights over all words in the vocabulary.  For each topic, we show the words with the 10 largest weights, along with a suggested description of the topic.}
\label{fig:topics}
\end{figure}
%\FloatBarrier

@@

\subsection*{Creating a low-dimensional embedding space}
Figures~7 and 8C in the main text display two-dimensional projections of the 100-dimensional topic trajectories for the video (7A, 8C), average recall (7B), and each individual's recall (7C, 8C).  We created these embeddings using the Uniform Manifold Approximation and Projection algorithm \citep[UMAP;][]{McInEtal18} called from our high-dimensional visualization and text analysis software, \texttt{HyperTools}~\citep{HeusEtal18a}.  An advantage of the UMAP algorithm over comparable manifold learning techniques (such as $t$-SNE) is that UMAP explicitly attempts to preserve the global structure of the data \citep{McInEtal18,BechEtal19} by constructing a space where distance on the manifold is standard Euclidean distance, with respect to the global coordinate system.  This was important in our use case, as we wanted to visualize both the evolving structure of the video and the spatial relationships between presented and recalled content.

@@

We performed a grid search over pre-specified values of these hyper-parameters, and optimized the manifold space for visualization based on two criteria: First, that the 2D embedding of the video trajectory should reflect its original 100-dimensional structure as faithfully as possible. Second, that the path traversed by the embedded video trajectory should intersect itself a minimal number of times.  The first criteria helps bolster the validity of visual intuitions about relationships between sections of video content, based on their locations in the embedding space.  The second criteria was motivated by the observed low off-diagonal values in the video trajectory's temporal correlation matrix (suggesting that the same topic-space coordinates should not be revisited; see Figure~2A in the main text) and the inability to accurately describe the consistency of participants' transitions through topic space at locations where those transitions frequently occurred in multiple, orthogonal directions (see \textit{Methods} for analysis details).  We formalized this optimization as the combination of hyper-parameter values that satisfied
\[
\argmax_{\left\{K, \gamma, \delta, \tau, \varphi~\in~E~\mathrm{|}~\Phi\left(K, \gamma, \delta, \tau, \varphi \right) \right\}} \left[\mathrm{corr}\left(\Upsilon\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right),~\Upsilon\left(\chi\right)\right)\right],~\mathrm{where}
\]
\[
\Phi\left(K, \gamma, \delta, \tau, \varphi \right) = \argmin_{K, \gamma, \delta, \tau, \varphi} \left[\Gamma\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right)\right],
\]
$\xi$ is the video's trajectory through the manifold space; $\chi$ is the original, 100-dimensional video trajectory; $\Upsilon$ is a function that computes a condensed matrix of pairwise distances between event vectors (computed using correlation distance in the original topic space and Euclidean distance in the manifold space); $\mathrm{corr}\left(\Upsilon\left(\xi\right), \Upsilon\left(\chi\right)\right)$ is the correlation between the sets of pairwise distances, and $\Gamma$ is the number of instances in which lines drawn between consecutive video event embeddings intersected each other.   The sets of hyper-parameter values we searched over comprised: $K \in \{10x \in \mathbb{Z}~|~10 < x <  22\} \cup \{161\}$ (a range roughly centered on half the total number of events, 161, in order to balance representations of local and global structure); $\gamma \in \{1, 3, 5, 7, 9\}$; $\delta \in \{0.1, 0.3, 0.5, 0.7, 0.9\}$; $\tau \in \{x \in \mathbb{Z}~|~0 < x < 101\}$; and $\varphi \in {S\choose3},~\mathrm{where}~S=\{\mathrm{video~events,~average~recall~events, individual~recall~events}\}$.  The optimal parameters (that yielded $\Phi=0$) were $K=170$, $\gamma=7$, $\delta=0.7$, $\tau=41$, with the order of sequence $\varphi$ as the average recall events, video events, and individuals' recall events, vertically concatenated, in order.
